---
title: "Clustering Analysis of scRNAseq data (subpopulation type)"
author: "Mayen Kalu(Inim)"
date: "2022-10-1"
output: html_document
---
# Load package dependencies
```{r}
library(Seurat)
library(tidyverse)
library(cowplot)
library(ggplot2)
library(ggrepel)
library(Matrix)
library(BiocManager)
library(GEOquery)
```

# Load datasets into R
# This is a scRNAseq obtained from NCBI GEO, published by Xie et al., 2021
```{r}
BC.raw <- read.csv(file = "C:/Users/Mayen/Documents/RBC_scRNAseq/clustering_analysis_rbc/GSE149938_bm/umi_matrix.csv")
BC.data <- Matrix(t(as.matrix(BC.raw)))
colnames(BC.data) %>%  head
dim(BC.data)
BC.meta <- GEOquery::getGEO("GSE149938", destdir = "C:/Users/Mayen/Documents/RBC_scRNAseq/clustering_analysis_rbc/GSE149938_bm/")
BC.meta
```

# Load merged data into a seurat object
```{r}
BC <- CreateSeuratObject(counts = BC.data, 
                         project = "RBC", 
                         min.cells = 0,
                         min.features = 0)
BC
VlnPlot(BC, features = "nFeature_RNA")
```

# Subset the data
```{r}
EC <- subset(BC, idents = 'ery')
EC
VlnPlot(EC, features = "nFeature_RNA")
```

# QC Metrics
# % MT
```{r}
EC[['percent.mt']] <- PercentageFeatureSet(EC, pattern = '^MT-')
EC[["percent.rbp"]] <- PercentageFeatureSet(EC, pattern = "^RP[SL]")
head(EC@meta.data, 3)
```

# Visualize QC metrics 
# as violin plot
```{r}
VlnPlot(EC, features = c('nFeature_RNA', 'nCount_RNA', 'percent.mt', 'percent.rbp'), ncol = 4)
```

# as ridge plot
```{r}
RidgePlot(EC, features = c('nFeature_RNA', 'nCount_RNA', 'percent.mt', 'percent.rbp'), ncol = 1)
```

# To visualize feature-feature relationships, FeatureScatter is used
```{r}
plot1 <- FeatureScatter(EC, feature1 = "nCount_RNA", feature2 = "percent.mt") 
plot2 <- FeatureScatter(EC, feature1 = "nCount_RNA", feature2 = "percent.rbp")
plot3 <- FeatureScatter(EC, feature1 = "nCount_RNA", feature2 = "nFeature_RNA")
plot1 + plot2 + plot3
```

# SCTransform
```{r}
EC <- SCTransform(EC, vars.to.regress = "percent.mt", verbose = FALSE)
```

# Visualizing the top 10 of highly variable features
```{r}
vf10 <- head(VariableFeatures(EC), 10)
vf10
plot4 <- VariableFeaturePlot(EC)
plot5 <- LabelPoints(plot = plot4, points = vf10, repel = TRUE)
plot4
plot5
```

# PCA analysis
```{r}
EC <- RunPCA(EC)
print(EC[["pca"]], dims = 1:20, nfeatures = 10)
VizDimLoadings(EC, dims = 1:2, reduction = 'pca')
DimPlot(EC, reduction = 'pca')
DimHeatmap(EC, dims = 1:6, cells = 500, balanced = TRUE)
```

# Determine dimensionality of the data
```{r}
ElbowPlot(EC)
```

# Clustering Analysis
# Determine the K-nearest neighbor graph
```{r}
EC <- FindNeighbors(EC, dims = 1:11)
```

# Determine the clusters for various resolutions  
```{r}
EC <- FindClusters(EC)
View(EC@meta.data)
head(Idents(EC), 5) 
```

# Run non-linear dimensional reduction (UMAP)
```{r}
EC <- RunUMAP(EC, reduction = "pca", dims = 1:11)
DimPlot(EC, reduction = "umap", label = TRUE, label.size = 4)
```

# Extract identity and sample information from seurat object to determine the number of cells per cluster per sample
```{r}
n_cells <- FetchData(EC, 
                     vars = c("ident", "orig.ident")) %>%
  dplyr::count(ident, orig.ident) %>%
  tidyr::spread(ident, n)
View(n_cells)
```

# Visualize metrics in EC@meta.data
```{r}
metrics <-  c("nCount_SCT", "nFeature_SCT", "percent.mt", "percent.rbp")
FeaturePlot(EC, 
            reduction = "umap",
            features = metrics,
            pt.size = 0.4, 
            sort.cell = TRUE,
            min.cutoff = 'q10',
            label = FALSE)
```

